---
title: "Introduction to ClusTorus"
author: "Seungki Hong"
date: "1/21/2021"
output: html_document
vignette: >
  %\VignetteIndexEntry{Introduction to ClusTorus}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

ClusTorus is a package for clustering angular data, especially for protein structure data. ClusTorus provides various clustering algorithms designed with conformal prediction framework, which can deal with the outliers. The package suggests various methods for fitting the algorithms, and some of them will be introduced soon. The package also provides some simple tools for handling angluar data, such as angular subtraction, computing angular distance, etc. Now, check how to use ClusTorus briefly.

## Data Loading and Handling

ClusTorus provides a protein structure data of SARS-CoV-2, which provoked the recent pandemic situation with COVID-19. With this data, we will introduce how to load the data and how to use data handling tools of ClusTorus. Since SARS-CoV-2 data is originally extracted from protein data bank, it contains various information for describing protein structure. For our purpose, we will only use `SARS_CoV_2$tbl`, which only contains angluar information of the data. Moreover, for simplicity, we will only use backbone-chain angles, that is, $\phi$ and $\psi$ columns.

```{r, message = FALSE, warning = FALSE}
library(ClusTorus)
library(tidyverse)

data <- SARS_CoV_2$tbl
data <- data.frame(data, id = rownames(data)) %>%
  separate(id,into = c(',','position','type','rest'))
data <- data %>% select(-",")

data <- data[(data$type == "B"), ]
data <- na.omit(data[, 1:2])
head(data)
```

To use the functions of ClusTorus, we need to convert the data to be on radian scale, especially on $[0, 2\pi)$. 
```{r}
data <- data / 180 * pi
data <- on.torus(data)
ggplot(data = as.data.frame(data), mapping = aes(x = phi, y = psi)) + geom_point()
```


`on.torus` converts the radian scale data to be on $[0, 2\pi)^d$, where $d$ means the number of dimension. In this case, $d = 2$ and thus the data is converted to be on $[0, 2\pi)^2$.

## Clustering with Various Options

Now, we are ready to implement clustering algorithms to the data. ClusTorus provides various options for clustering, but we will provide only several cases: "kde", "mixture - axis aligned", and "kmeans - general". On the other hand, we need to choose hyperparameters: the number of modes or ellipsoids $J$ and the significance level $\alpha$. Before choosing the hyperparameter, we will implement the model fitting function `icp.torus.score` with various hyperparameter options, first.

```{r, warning = FALSE}
set.seed(2021)

Jvec <- 3:35
l <- list()

for (j in Jvec){
  l[[j]] <- icp.torus.score(as.matrix(data),
                            method = "all",
                            mixturefitmethod = "axis-aligned",
                            kmeansfitmethod = "general",
                            param = list(J = j, concentration = 25), 
                            verbose = FALSE)
}

```


The list `l` contains the `icp.torus` objects, which consist of fitted parameters for generating clusters, by varying the hypterparameter $J$. That is, These objects are optimally fitted ingredients for generating clusters for given $J$. By specifying the significance level, we can get the clusters. But, how to generate optimal clusters? One may think that the hyperparameter which generates the cluster of the minimum volume/area will be the optimum for given significance level. For dealing with such volume/area based hyperparameter searching, ClusTorus provides `icp.torus.eval` function, which generates boolean matrices for indicating the inclusion of pre-specified evaluation points for given significance level(s). You need to know that the result also changes by varying $\alpha$. Thus, we need to carefully choose $\alpha$ for better performance under pre-specified hyperparameter-choosing criterion. But, for simplicity, we will assume that the significance level $\alpha$ is given with $0.1$. Then, we can simply get the results by using pre-obtained `icp.torus` objects as below:

```{r}
alpha <- 0.1
out <- data.frame()

for (j in Jvec){
  a<-icp.torus.eval(l[[j]], level = alpha, eval.point = grid.torus())
  mu_kde <- sum(a$Chat_kde[,1])
  mu_e <- sum(a$Chat_e[,1])
  mu_kmeans <- sum(a$Chat_kmeans[,1])
  out <- rbind(out, data.frame(J = j, mu_kde = mu_kde, mu_e = mu_e, mu_kmeans = mu_kmeans))
}

head(out)
```

We count the number of points included in the generated clusters. In this 2-dimensional case, `grid.torus()` generates the 10,000 grid points in default. That is, `icp.torus.eval` generates the matrices for indicating whether each grid point is included in the clusters. By counting the included points, we can approximately estimate the volume of clusters. `mu_kde` shows the volume of cluster generated by option `kde`, `mu_e` by option `mixture` with `axis-aligned`, and `mu_kmeans` by option `kmeans` with `general`. Now, with "minimum volume" criterion, choose optimal $J$ for each method, and generate the clusters.

```{r}
J_kde <- out$J[which.min(out$mu_kde)]
J_e <- out$J[which.min(out$mu_e)]
J_kmeans <- out$J[which.min(out$mu_kmeans)]

icp.torus.kde <- l[[J_kde]]
icp.torus.mix <- l[[J_e]]
icp.torus.kmeans <- l[[J_kmeans]]

c_mix <- cluster.assign.torus(data, icp.torus.mix, level = alpha)
c_kmeans <- cluster.assign.torus(data, icp.torus.kmeans, level = alpha)
```

If there are some difficulties for evaluating such volume or for implementing your hyperparameter-choosing criterion, ClusTorus alternatively suggests the function `hyperparam.torus`, which suggests two hyperparameters $J$ and $\alpha$. For the choice of $J$, there are several options such as "elbow", "risk", "AIC" and "BIC". The option "elbow" is actually the volume based criterion, which is already described above. "risk" is the criterion based on the sum of the conformity scores which are evaluated under the inductive conformal prediction framework, that is, evaluated with the function `icp.torus.score`. "AIC" and "BIC" are well-known information criteria, proposed by Akaike (1974) and Schwarz (1978) respectively. These options can be designated with the argument `option`. The function `hyperparam.torus` generates `data.frame` objects for $J$ versus the evaluated criterion and for $\alpha$ versus the number of resultant clusters. It also generates `icp.torus` object based on the selected $J$.

```{r}
output <- hyperparam.torus(data, icp.torus.objects = l, option = "risk")
output$IC.results
output$alpha.results
output$optim$hyperparam
```


## Generating Clusters and Visualization of Clustering Results
With `cluster.assign.torus`, we can generate the cluster for option `mixture` and `kmeans`, and for each data point, the label of cluster is assigned. Moreover, for the cases of `mixture` and `kmeans`, we can check how the cluster is generated directly as below:

```{r}
c_mix$mixture$plot
c_kmeans$kmeans$plot
```

Also, with the assigned cluster labels, we can visualize the clusters with colored data points; 
```{r}
library(cowplot)
result.dat.mix <- data.frame(data, membership = as.factor(c_mix$mixture$cluster.id.outlier)) %>%
  mutate(membership = ifelse(membership == max(c_mix$mixture$cluster.id.outlier), "out", membership))
result.dat.kmeans <- data.frame(data, membership = as.factor(c_kmeans$kmeans$cluster.id.outlier)) %>%
  mutate(membership = ifelse(membership == max(c_kmeans$kmeans$cluster.id.outlier), "out", membership))

g1 <- ggplot(data = result.dat.mix, aes(phi,psi, color = membership)) + geom_point() +
  scale_x_continuous(breaks = c(0,1,2,3,4)*pi/2, labels = c("0","pi/2","pi","3pi/2","2pi"), limits = c(0,2*pi))+
  scale_y_continuous(breaks = c(0,1,2,3,4)*pi/2, labels = c("0","pi/2","pi","3pi/2","2pi"), limits = c(0,2*pi))
g2 <- ggplot(data = result.dat.kmeans, aes(phi,psi, color = membership)) + geom_point() +
  scale_x_continuous(breaks = c(0,1,2,3,4)*pi/2, labels = c("0","pi/2","pi","3pi/2","2pi"), limits = c(0,2*pi))+
  scale_y_continuous(breaks = c(0,1,2,3,4)*pi/2, labels = c("0","pi/2","pi","3pi/2","2pi"), limits = c(0,2*pi))

plot_grid(g1, g2)
```

On the other hand, to check the clustering result of method `kde`, we can visualize the results as below:
```{r}
ia <- icp.torus.eval(icp.torus.kde, level = alpha, eval.point = grid.torus())
b <- data.frame(ia$eval.point, ia$Chat_kde == 1)
colnames(b) <- c("phi","psi","C_kde")

g0 <- ggplot() + geom_contour(aes(phi, psi, z = ifelse(C_kde,1,0)), data = b, size = 1,lineend = "round" ) +
  geom_point(mapping = aes(x,y), data = data.frame(x = data[,1],y =data[,2])) +
  scale_x_continuous(breaks = c(0,1,2,3,4)*pi/2, labels = c("0","pi/2","pi","3pi/2","2pi"), limits = c(0,2*pi))+
  scale_y_continuous(breaks = c(0,1,2,3,4)*pi/2, labels = c("0","pi/2","pi","3pi/2","2pi"), limits = c(0,2*pi))
g0
```

If you clearly follow the examples, you may need the detailed manual. Check the package manual and see the details. 




